---
title: 'Forecasting House Price of Wards in the Borough of Southwark '
author: "Yiru Xu"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Preparation

```{r load_packages,include=FALSE}
library(sf)
library(sp)
library(dplyr)
library(tmap)
library(ggplot2)
library(tseries)
library(forecast)
source('starima_package.R')
library(spdep)
library(knitr)
library(reshape2)
library(rgdal)
```

```{r feature_engineering, include=FALSE}
# read the data
london_house_price = read.csv('data/mhp_london_ward.csv')
london_ward = st_read("data/London_Ward.shp")
# remove unnecessary rows and columns
london_house_price = london_house_price[-c(1,2,3,4),]
london_ward = london_ward[, -c(5,6)]
# change column names for the two data sets
colnames(london_house_price) = london_house_price[1,]
colnames(london_ward)[2] = london_house_price[1,3]
# remove columns of 1995Dec, 2022Mar and 2022Jun 
london_house_price = london_house_price[-1,-c(5,110:113)]
# filter southwark wards and delete unnecessary columns
southwark_ward = filter(london_ward, DISTRICT == 'Southwark')
# filter southwark wards from the house price data
southwark_house_price = 
  london_house_price[london_house_price$`Ward code` %in% southwark_ward$`Ward code`, ]
southwark_house_price = southwark_house_price[,-c(1,2)]
# replace ":" into "0,000" - easier for transforming
southwark_house_price[southwark_house_price == ':'] = '0,000'
# change the house price data type into numeric
southwark_house_price[, 3:ncol(southwark_house_price)] = 
  apply(southwark_house_price[, 3:ncol(southwark_house_price)], 
        2, function(x) as.numeric(gsub(",", "", x)))
# aggregate house prices to wards geometry data
sk_ward_hp = inner_join(southwark_ward, 
                        southwark_house_price[, c(1,3:ncol(southwark_house_price))], by = c("Ward code" = "Ward code"))
sk_ward_hp = sk_ward_hp[ , -c(2:4)]
# change the column names 
months = gsub('Year ending (\\w{3}) (\\d{4})', '\\1', 
              colnames(sk_ward_hp)[2:(ncol(sk_ward_hp)-1)])
years = gsub('Year ending (\\w{3}) (\\d{4})', '\\2', 
             colnames(sk_ward_hp)[2:(ncol(sk_ward_hp)-1)])
new_colnames = paste(months, as.character(years), sep = '-')
colnames(sk_ward_hp)[2:(ncol(sk_ward_hp)-1)] = new_colnames
```

```{r convert_data, include=TRUE}
# create a matrix for easier analysis
sk_ward_hp_matrix = data.matrix(sk_ward_hp[2:(ncol(sk_ward_hp)-1)])
sk_ward_hp_matrix = sk_ward_hp_matrix[, -105]
# calculate the mean of each year
# create an empty matrix to store the means
sk_ward_yhp_matrix = matrix(0, nrow = nrow(sk_ward_hp_matrix), ncol = 26)
# Loop over each row of the matrix and calculate the mean of every year
for (i in 1:nrow(sk_ward_hp_matrix)) {
  for (j in 1:26) {
    sk_ward_yhp_matrix[i, j] = mean(sk_ward_hp_matrix[i, ((j-1)*4+1):(j*4)])
  }
}
colnames(sk_ward_yhp_matrix) = c(1996:2021)
# change yearly data to dataframe
sk_ward_yhp = data.frame(sk_ward_yhp_matrix)
colnames(sk_ward_yhp) = c(1996:2021)
# add a ward code column for aggregation
sk_ward_yhp[,27] = southwark_house_price['Ward code']
# aggregate house prices to wards geometry data
sk_ward_yhp = inner_join(southwark_ward, 
                        sk_ward_yhp, by = c("Ward code" = "Ward code"))
sk_ward_yhp = sk_ward_yhp[ , -c(2:4)]
```

```{r write_shp, include=TRUE}
# Convert the seasonal data frame to an sf object
southwark_hp = st_as_sf(sk_ward_hp, crs = 27700)
# Write the sf object to a shapefile
st_write(southwark_hp, "southwark_hp.shp")
# Convert the yearl data frame to an sf object
southwark_yhp = st_as_sf(sk_ward_yhp, crs = 27700)
# Write the sf object to a shapefile
st_write(southwark_yhp, "southwark_yhp.shp")
```

## ESTDA

#### Non Spatio-temporal Characteristics

```{r data_distribution, include=True}
# statistic summary of southwark and dulwich village
print(paste("Southwark mean house price:",mean(sk_ward_hp_matrix)))
print(paste("Dulwich village mean house price:",mean(sk_ward_hp_matrix[3,])))
print(paste("Southwark median house price:",median(sk_ward_hp_matrix)))
print(paste("Dulwich village median house price:",median(sk_ward_hp_matrix[3,])))
print(paste("Southwark house price standard deviation:",sd(sk_ward_hp_matrix)))
print(paste("Dulwich village house price standard deviation:",median(sk_ward_hp_matrix[3,])))
print(paste("Southwark house price(min):",min(sk_ward_hp_matrix)))
print(paste("Dulwich village house price(min):",min(sk_ward_hp_matrix[3,])))
print(paste("Southwark house price(max):",max(sk_ward_hp_matrix)))
print(paste("Dulwich village house price(max):",max(sk_ward_hp_matrix[3,])))
# distribution of all the wards
par(mfrow=c(1,2))
hist(sk_ward_hp_matrix, main = 'Southwark', xlab = 'House Price')
mu = mean(sk_ward_hp_matrix)
abline(v=mu, col='red')
# distribution of Dulwich Village
hist(sk_ward_hp_matrix[3,], main = 'Dulwich Village', xlab = 'House Price')
mu = mean(sk_ward_hp_matrix[3,])
abline(v=mu, col='red')
```

#### Temporal Characteristics

```{r temporal_pattern,include = TRUE}
# Reshape data from wide to long format
sk_ward_long = tidyr::pivot_longer(sk_ward_hp, cols = 2:(ncol(sk_ward_hp)-1), 
                             names_to = "Date", values_to = "Price")
colnames(sk_ward_long)[1] = 'Ward'
# change the Date column to 'Date" class for visualisation   
sk_ward_long$Date = lubridate::dmy(paste0('30-',sk_ward_long$Date))
# plot the temporal pattern
ggplot(sk_ward_long,
       aes(x = Date, y = Price, group = Ward, color= Ward)) + 
  geom_line() + 
  xlab('Time') +
  ylab("House Price") + 
  scale_x_date(limits = c(min(sk_ward_long$Date), max(sk_ward_long$Date)),
               date_labels = "%Y", date_breaks = "1 year") + 
  ggtitle('House Price of Wards in Southwark (1996-2021)')+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  theme(plot.title = element_text(hjust = 0.5))
# only for dulwich Village
# change the data matrix into time series object
dulvil_hp_ts = ts(sk_ward_hp_matrix[3,], start = c(1996, 1), end = c(2021, 4), frequency = 4)
plot(dulvil_hp_ts, type = 'l', xlab = 'Time', ylab = 'House Price', main = 'House Price in Dulwich Village from 1996 to 2021')
```

```{r temproal_autocorrelation, include=TRUE}
# acf and pacf
par(mar = c(5, 4, 4, 5) + 0.1)
acf(dulvil_hp_ts, lag.max = 36, xlab = 'Lag', ylab = 'ACF', main = 'ACF of Seasonal House Price in Dulwich Village')
pacf(dulvil_hp_ts, lag.max = 36, xlab = 'Lag', ylab = 'PACF', main = 'PACF of Seasonal House Price in Dulwich Village')
# scatter plot
par(mfrow=c(2,4))
lag.plot(dulvil_hp_ts, lags = 16, do.lines = FALSE)
```

#### Spatial Characteristics

```{r mapping_data, include=TRUE}
southwark_yhp_shp = st_read('southwark_yhp.shp')
colnames(southwark_yhp_shp)[2:27]= c(1996:2021)
for (j in 2:(ncol(southwark_yhp_shp)-1)) {
  par(mfrow=c(4,6))
  map = tm_shape(southwark_yhp_shp)+
    tm_fill(colnames(southwark_yhp_shp)[j], style="jenks", palette="Purples")+
    tm_borders("white")+
    tm_compass(position=c("right","top"))+
    tm_scale_bar()
  print(map)
} 
```

```{r spatial_autocorrelation, include=TRUE}
# spatial autocorrelation in 26-year scale
southwark_hp_shp = st_read('southwark_hp.shp')
nb = poly2nb(southwark_hp_shp)
lw = nb2listw(nb)
sk_ward_hp_avg = rowMeans(sk_ward_hp_matrix)
moran.test(sk_ward_hp_avg, listw=lw1)
# spatial autocorrelation in 1-year scale
for (m in 2:(ncol(southwark_yhp_shp)-1)) {
  par(mfrow=c(4,6))
  nb1 = poly2nb(southwark_yhp_shp)
  lw1 = nb2listw(nb1)
  sk_ward_yhp_avg = sk_ward_hp_matrix[,m]
  print(moran.test(sk_ward_yhp_avg, listw=lw1))
}
```

## Data Decomposition and Differencing

```{r data_decomposition, include=TRUE}
decomp <- decompose(dulvil_hp_ts, type="multiplicative")
plot(decomp, type = 'l')
```

```{r differencing_acf_pacf, include = True}
dulvil_hp_df = diff(dulvil_hp_ts, differences = 1)
# Augmented Dickey-Fuller Test (ADF) for stationary
adf.test(dulvil_hp_df)
# differenced time series pattern
plot(dulvil_hp_df, type = 'l', xlab = 'Time', ylab = 'House Price', main = 'House Price in Dulwich Village from 1996 to 2021(Differenced)')
# check acf and pacf
par(mfrow = c(1,2))
par(mar = c(5, 4, 4, 5) + 0.1)
acf(dulvil_hp_df, lag.max = 40, xlab = 'Lag', ylab = 'ACF', main = 'ACF of Differenced House Price in Dulwich Village')
pacf(dulvil_hp_df, lag.max = 40, xlab = 'Lag', ylab = 'PACF', main = 'PACF of Differenced House Price in Dulwich Village')
```

```{r seasonal_differencing, include=TRUE}
# seasonal differencing
dulvil_hp_sdf = diff(dulvil_hp_df, differences = 1, lag = 8)
# comparison between differenced and seasonal-differenced data 
par(mar = c(5, 4, 4, 5) + 0.1)
par(mfrow = c(1,2))
# acf
acf(dulvil_hp_df, lag.max = 40, xlab = 'Lag', ylab = 'ACF', main = 'Differenced')
acf(dulvil_hp_sdf, lag.max = 40, xlab = 'Lag', ylab = 'ACF', main = 'Differenced + Seasonal Differenced')
# pacf
pacf(dulvil_hp_df, lag.max = 40, xlab = 'Lag', ylab = 'PACF', main = 'Differenced')
pacf(dulvil_hp_sdf, lag.max = 40, xlab = 'Lag', ylab = 'PACF', main = 'Differenced+ Seasonal Differenced')
par(mfrow = c(2,1))
plot(dulvil_hp_sdf, type = 'l', main = 'Differenced Time Series')
plot(dulvil_hp_df, type = 'l', main = 'Differenced + Seasonal Differenced Time Series')
# Augmented Dickey-Fuller Test (ADF) for stationary in residuals
random_no_na = na.omit(decomp$random)
adf.test(random_no_na)
```

## Parameter Estimation (ACF & PACF Analysis)

```{r order_selection, include=TRUE}
par(mar = c(5, 4, 4, 5) + 0.1)
acf(dulvil_hp_sdf, lag.max = 40, xlab = 'Lag', ylab = 'ACF', main = 'ACF of House Price in Dulwich Village (Differenced + Seasonal Differenced)')
pacf(dulvil_hp_sdf, lag.max = 40, xlab = 'Lag', ylab = 'ACF', main = 'PACF of House Price in Dulwich Village (Differenced + Seasonal Differenced)')
```

## Model Fitting

### ARIMA

Dividing the data into training data(1996-2015) and testing data(2016-2021).

```{r sarima_110110), include=TRUE}
fit_ar = arima(dulvil_hp_ts[1:84],order=c(1,1,0), seasonal=list(order=c(1,1,0)
,period=8))
print(fit_ar)
```

```{r diagnostic, include=TRUE}
# a diagnostic plot of the fitted model
tsdiag(fit_ar)
```

```{r sarima_with_MA, include=TRUE}
fit_ar = arima(dulvil_hp_ts[1:84],order=c(4,1,4), seasonal=list(order=c(1,1,0)
,period=8))
print(fit_ar)
```

```{r diagonostic, include=TRUE}
# a diagnostic plot of the fitted model
tsdiag(fit_ar)
```

```{r prediction, include=TRUE}
# define a function to calculate NRMSE for prediction
nrmse <- function(actual, predicted) {
  rmse <- sqrt(mean((actual - predicted)^2))
  range <- max(actual) - min(actual)
  return(rmse / range)
}
pre_ar = predict(fit_ar, n.ahead=20)
matplot(1:20,cbind(dulvil_hp_ts[85:104],pre_ar$pred),type="l",main=
"SARIMA prediction effect", ylab = 'Time')
NRMSE_fit = NRMSE(fit_ar$residuals, dulvil_hp_ts[1:84])
print(paste('fit_NRMSE:',NRMSE_fit))
NRMSE_pre = nrmse(dulvil_hp_ts[85:104],pre_ar$se)
print(paste('pre_NRMSE:',NRMSE_pre))
```

## STARIMA

## Building spatial weight matrix

```{r spatial_weight_matrix, include=TRUE}
Wmat <- listw2mat(lw)
```

## **Space-time Autocorrelation and partial autocorrelation analysis**

```{r acf_pacf, include=TRUE}
sk_ward_hp_tm = t(sk_ward_hp_matrix)
stacf(sk_ward_hp_tm, Wmat, 40)
stpacf(sk_ward_hp_tm, Wmat, 40)
```

```{r differencing, include=TRUE}
# differencing the matrix
sk_ward_hp_tmdf = diff(sk_ward_hp_tm, differences = 1)
# change the spatial weight matrix to fit the differenced matrix
stacf(sk_ward_hp_tmdf, Wmat, 40)
stpacf(sk_ward_hp_tmdf, Wmat, 40)
```

## Model Fitting

```{r starima, include=TRUE}
W_fit = list(w1=Wmat)
fit_star = starima_fit(sk_ward_hp_tm[1:84,],W_fit,p=1,d=1,q=0)
```

```{r diagnostic, include=TRUE}
stacf(fit_star$RES,Wmat,40)
```

```{r residuals_hist,include=TRUE}
hist(fit_star$RES[,3])
```

## Prediction

```{r prediction, include=TRUE}
pre_star = starima_pre(sk_ward_hp_tm[84:104,], model = fit_star)
matplot(1:20,cbind(sk_ward_hp_tm[85:104,3],pre_star$PRE[,3]),type="l", main = 'STARIMA prediction effect')
NRMSE_star_fit = NRMSE(fit_star$RES, dulvil_hp_ts[1:84])
print(paste('stfit_NRMSE: ',NRMSE_star_fit))
NRMSE_star_pre = nrmse(dulvil_hp_ts[85:104],pre_star$PRE)
print(paste('stpre_NRMSE: ',NRMSE_star_pre))
```

## 
